---
title: "HW04 Part 2: Graph some biological data"
author: "Molly Brakhane"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r}
library(tidyverse)
```

# Plot 1
```{r}
data(trees)
```
```{r}
?trees
```

```{r}
ggplot(data = trees) +
  geom_point(mapping = aes(x = Height, y = Girth))
```

There is positive trend in girth and height, meaning they increase together.

# Plot 2
```{r}
year <- 1821:1934
```
```{r}
data(lynx)
?lynx
class(lynx)
pelts <- as.vector(lynx)
```
```{r}
lynx_pelts <- data.frame(year, pelts)
ggplot(data = lynx_pelts)+
  geom_line(mapping = aes(x = year, y = pelts, color = "maroon"))
```
Whenever there is a lot of pelts, it decreases after until the population grows again. Then the number of pelts goes back up,
# Plot 3
```{r}
data(iris)
?iris
```
```{r}
ggplot(data = iris) +
  geom_point(mapping = aes(x = Petal.Length, y = Petal.Width, color = Species, shape = Species), size = 2) +
  labs(x = "Petal Length", y = "Petal Width")
```
Virginica is the largest, and the versicolor is second. The smallest is setosa.
# Plot 4 & 5
```{r}
?geom_violin
?msleep
ggplot(data = msleep, mapping = aes(x = vore, y = sleep_rem)) +
  geom_violin(fill = "gray70")
```
Carnivores have very little REM sleep. Insects have more. Most herbivores also get very litte. Gray 30 is much darker than gray 70.
# Plot 7
```{r}
ggplot(data = msleep) +
  geom_boxplot(mapping = aes(x = order, y = sleep_total)) +
  coord_flip()
```

# Plot 8
```{r}
ggplot(data = msleep, mapping = aes(x = conservation, y = awake, color = conservation)) +
  geom_boxplot()+
  geom_point() +
  geom_jitter() +
  coord_flip() +
  labs(x = "Awake", y = "Conservation", color = "Conservation")
```

# Plot 9 & 10
```{r}
ggplot(data = msleep) +
  geom_point(mapping = aes(x = awake, y = bodywt, color = vore))
ggplot(data = msleep)+
  geom_point(mapping = aes(x = awake, y = brainwt, color = conservation)) +
  facet_wrap(.~vore, ncol = 2) 
```
Herbivores are awake more than other animals. Animas with more brain wight also sem to stay awake longer than other animals.Herbivores are larger in general.
