---
title: "HW04 Part 1: Complete Chapter 3"
author: "Molly Brakhane"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Chapter 3

- Change "your name" in the YAML header above to your name.

### Section 3.1

Load the tidyverse packages. I've entered the first code chunk for you. **You must enter all subsequent code chunks and run the code.** 

```{r}
library("tidyverse")
```

If you get an error, then you did not install the `tidyverse` package like you were supposed to in the previous assignment. Go back to that assignment and complete that step. (This should not be the case, though.)

### Section 3.2: First steps

Enter your code chunks for Section 3.2 here.

Get in the habit of entering a brief description of what each chunk of code does. Enter this above the chunk, like I did in Section 3.1.

```{r}
mpg
```
Opening the data frame

```{r}
?mpg

```

#### Section 3.2 Questions

Answer the questions *completely.* Some answers will require you to write code in a code chunk and also to type text before or after the chunk. Other answers will require only code or only text. 

**1:** Run ggplot(data = mpg). What do you see?
```{r}
ggplot(data = mpg)
```
I see a gray box? I hope thats what I'm supposed to see. 

**2:** How many rows are in `mpg`? How many columns? **Hint:** Use the `dim()` function. Type `dim()` in the console to learn what `dim()` does.
```{r}
dim(mpg)
```
234 rows and 11 columns.

**3:** What does the `drv` variable describe? Read the help for `?mpg` to find out.
drv describes whether the car has front, rear, or 4 wheel drive.

**4:** Make a scatterplot of `hwy` vs `cyl`.
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = cyl, y = hwy))
```


**5:** What happens if you make a scatterplot of `class` vs `drv`? Why is the plot not useful?
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = class, y = drv))
```
The scatterplot appears random and is not a good way to display this data. It should be disiplayed another way.

### Section 3.3: Aesthetic mappings
```{r}
ggplot(data=mpg) +
  geom_point(mapping = aes(x=displ, y = hwy, color = class))
```
Map an aesthetic to a varaiable, associate the aesthetic to the name of the varaible inside aes().

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, size = class))
  
```
Size is ordered, class is not. not a good pair.

```{r}
ggplot(data=mpg) + 
  geom_point(mapping = aes(x=displ, y=hwy, alpha = class))
```
```{r}
ggplot(data=mpg) + 
  geom_point(mapping = aes(x=displ, y=hwy, shape = class))
```
Will only plot 6 shapes at a time!!

```{r}
ggplot(data=mpg) + 
  geom_point(mapping = aes(x=displ, y=hwy), color = "blue")
```


#### Section 3.3 questions

**1:** What’s gone wrong with this code? Why are the points not blue?
```{r}
ggplot(data=mpg) + 
  geom_point(mapping = aes(x=displ, y=hwy, color = "blue"))
```
The addition of the color wasn't added outside the aes(), but within, so "blue" was used in the legend and the default color was added.

**2:** Which variables in `mpg` are categorical? Which variables are continuous? (*Hint:* type `?mpg` to read the documentation for the dataset). How can you see this information when you run mpg?
Class, Drv, abd manufacturer model are categorical. They specify only a few possibilities. Continous variables, like year, cty, and hwy can by anything within.

**3:** Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs. continuous variables? **Note:** You need several code chunks here to answer this question.
```{r}
ggplot(data=mpg) + 
  geom_point(mapping = aes(x=class, y=hwy, shape = year))
```
 It won't map to shape at all.
```{r}
ggplot(data=mpg) + 
  geom_point(mapping = aes(x=displ, y=hwy, color = year))
```
 Color will it a gradient, sort of like the alpha.
 
```{r}
ggplot(data=mpg) + 
  geom_point(mapping = aes(x=displ, y=hwy, size = year))
```
 Mapping it to size is not very helpful, it narrowed the data available down.

**4:** What happens if you map the same variable to multiple aesthetics?
```{r}
ggplot(data=mpg) + 
  geom_point(mapping = aes(x=class, y=year, alpha = year))
```
it gives a useless table.

**5:** What does the stroke aesthetic do? What shapes does it work with? (*Hint:* use `?geom_point`)

```{r}
?geom_point
```
Stroke will make the weight of the shape borders stronger or lighter.

**6:** What happens if you map an aesthetic to something other than a variable name, like `aes(colour = displ < 5)`?
```{r}
ggplot(data=mpg) + 
  geom_point(mapping = aes(x = hwy, y = cyl, color=displ<5))
```
It will determine and color-code the data if it is true or not.

### Section 3.5: Facets

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x =displ, y = hwy)) +
  facet_wrap(~class, nrow = 2)
```
A facet will display your data in subplots to show all of your data.

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x =displ, y = hwy)) +
  facet_grid(drv ~ cyl)
```


#### Section 3.5 questions

**1:** What happens if you facet on a continuous variable? **Note:** Write a prediction, then make and run your code chunk.
I don't think it will make any sense because the facet will end up being too long to be meaningful.
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = year, y = class)) +
  facet_wrap(~hwy, nrow = 3)
  
```




**2:** What do the empty cells in plot with `facet_grid(drv ~ cyl)` mean? How do they relate to this plot?

There are no cars that have that number of cylinders and that kind of driving pattern. This means that those don't exist, which is definitely important to know.

**3:** What plots does the following code make? What does `.` do?

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(. ~ cyl)
```
Using '.' we can facet it to one variable instead of two.

**4:** Take the first faceted plot below. What are the advantages to using faceting instead of the colour aesthetic? What are the disadvantages? How might the balance change if you had a larger dataset?
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)

```
Faceting separates the data to a more readable version. If there was a larger data set, the faceting might get a little complicated but the color aesthetic would also get distracting. Color is more visually appealing and better for more than 7 classes.


**5:** Read `?facet_wrap`. What does `nrow` do? What does `ncol` do? What other options control the layout of the individual panels? Why doesn’t `facet_grid()` have nrow and ncol argument?
```{r}
?facet_wrap
```
Ncol and nrow decide the number of rows and columns.

**6:** When using `facet_grid()` you should usually put the variable with more unique levels in the columns. Why?

### Section 3.6: Geometric objects
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy))
```
```{r}
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy))
```

```{r}
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv))
```
```{r}
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy, color = drv))
```
```{r}
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy, group = drv))
```
```{r}
ggplot(data = mpg) +
  geom_smooth(
    mapping = aes(x = displ, y = hwy, color = drv),
    show.legend = FALSE
    )
```
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  geom_smooth(mapping = aes(x = displ, y = hwy))
```
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y  = hwy)) +
  geom_point(mapping = aes(color = class)) +
  geom_smooth(data = filter(mpg, class == "subcompact"), se = FALSE)
```

#### Section 3.6.1 Questions

**1:** What geom would you use to draw a line chart? A boxplot? A histogram? An area chart? **Note:** You will have to make some types of these charts in Part 2 of this assignment.
Line chart would be geom_line
Boxplot chart would be geom_boxplot
Histogram would be geom_histogram
An area chart would be geom_area.


**2:** Run this code in your head and predict what the output will look like. Then, run the code in R and check your predictions.
x axis is displacement, y is highway mpg, and the dots will be determined by what kind of drive the car has. There will be lines too.
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```


**3:** What does `show.legend = FALSE` do? What happens if you remove it?
**Note:** Skip the "Why do you think I used it earlier in the chapter?" question.
Show.legend = FALSE prevents the legend from being shown. Otherwise, the legend will be shown.

**4:** What does the `se` argument to `geom_smooth()` do?
Se = false prevents the shadow of the line, making a single line instead.

**5:** Will these two graphs look different? Why/why not?
No, they will be the same. They are making the same graph, just written slightly differently.

**6:** Recreate the R code necessary to generate the following graphs. **Note:** Enter a separate code chunk for each graph, numbered 1 to 6, starting row-wise at upper left. 

1: upper left.
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point() +
  geom_smooth(se = FALSE)
```


2: upper right.
```{r}
ggplot() +
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy, group = drv), se = FALSE)
```


3: middle left.
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) +
  geom_point() +
  geom_smooth(se = FALSE)
```


4: middle right.
z
```{r}
ggplot() +
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy, color = drv))+
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy), se = FALSE)
```

5: lower left.
```{r}
ggplot() +
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy, color = drv))+
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy, linetype = drv), se = FALSE)
```


6: lower right. *Hint:* Read the help file for `geom_point`. Use `shape = 21`, `stroke = 2`, and `size = 3`.
```{r}
ggplot() +
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy, color = drv), stroke = 2, size = 3, shape = 21)
```

```{r}
?geom_point
```

### Section 3.7: Statistical transformations.

**Skip this section.** We may return to it later.


### Section 3.8 Position adjustments
```{r}
ggplot(data=diamonds) + 
  geom_bar(mapping = aes(x = cut, color = cut))
ggplot(data=diamonds) +
  geom_bar(mapping = aes(x = cut, fill = cut))
```
```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = clarity))
```
```{r}
ggplot(data = diamonds, mapping = aes(x = cut, fill = clarity)) +
  geom_bar(alpha = 1/5, position = "identity")
ggplot(data = diamonds, mapping = aes(x = cut, color = clarity)) +
  geom_bar(fill = NA, position = "identity")
```
```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x=cut, fill = clarity), position = "fill")
```
```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "dodge")
```
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy), position = "jitter")
```


#### Section 3.8 Questions

**1:** What is the problem with this plot? How could you improve it. *Hint:* Remember the size of the `mpg` data set from earlier in this exercise. Does the number of points seem correct? **Note:** Write the code to improve it. 
```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point(position = "jitter")
```


**2:** What parameters to `geom_jitter()` control the amount of jittering?
```{r}
?geom_jitter
```
width and height

**3:** Compare and contracts `geom_jitter()` with `geom_count()`.
```{r}
ggplot(data = mpg) +
  geom_jitter(mapping = aes(x = displ, y = hwy))
ggplot(data = mpg) +
  geom_count(mapping = aes(x = displ, y = hwy))
```
Count adds all data points in a spot, while jitter separates them to make them more visible.

**4:** What is the default position adjustement for `geom_boxplot()`? Create a visualization of the `mpg` dataset that demonstrates it.
```{r}
ggplot(data = mpg) +
  geom_boxplot(mapping = aes(x = displ, y = hwy, group = class))
```


### Section 3.9: Coordinate systems

**Note:** Read about and do only the code chunk associated with `coord_flip()`. `coord_flip()` is a great layer to use when you have categorical data with longish names as it makes the text easier to read.
```{r}
ggplot(data = mpg, mapping = aes(x = class, y =hwy))+
  geom_boxplot()
ggplot(data = mpg, mapping = aes(x=class, y = hwy)) +
  geom_boxplot() +
  coord_flip()
```
 

### Section 3.10: The layered grammar of graphics

Just read this section as a contextual review of everything you just learned.


