---
title: "R Notebook"
output: html_notebook
---
#Activity 5 - Data Manipulation Sampler but i didn't use gapminder bc i forgot until 10 minutes before due
#Establish data with data.frame. C# creates columns.
```{r}
A <- data.frame(
       c1 = c('A', 'B', 'C', 'D', 'E'),
       c2 = c(1, 3, 2, 2, 1),
       c3 = c(5, 3, 9, 6, 1))

B <- A %>%
  group_by(c1,c2) %>%
  summarize(Nrows=n(),
            SumOfCol=sum(c2 + c3))

```

```{r}
A
```

```{r}
B
```
#Creating a table without reshape2
```{r}
C <-with(A, table(c1, c2))
C
```
#using reshape2 to make a dataframe
```{r}
C_alt<-dcast(A, c1~c3,value.var="c2",fun.aggregate=length)
C_alt
```
```{r}
library(reshape2)
```
###oH THANK GOODNESS I FINALLY GOT IT TO UPLOAD
###STILL NO EXPLANATION ON KABLE

#Reformatting a table
```{r}
D <- melt(C, id="c1") %>% 
  arrange(c1)
D
```

#use merge to combine two tables
```{r}
E <- data.frame(
  c1 = c("A", "B", "C"),
  c2 = c(1, 2, 5)
)

F <- data.frame(
  c1 = c("A", "B", "B", "B", "C", "A"),
  c3 = c("a", "a", "a", "b", "c", "b")
)

G <-merge(F,E)
```
```{r}
E

F

G
```

#Filter creates a subset of the data
```{r}
H <- filter(G, c1=="A")
H
```

#Use transpose to re-organize the data
```{r}
H_transpose <- data.frame(t(H))
H_transpose
```

```{r}
a_arranged <- arrange(A,c1,c2,c3)
a_arranged
```

#Joining

```{r}
gapminder
mtcars
merge(gapminder, mtcars)
```

```{r}
Wowee <- merge(gapminder, mtcars)
Wowee
filter(Wowee, gear==4, year==1952)
filter(Wowee, continent=="Asia", lifeExp>45, year>2000)
```

```{r}
match("Continent", gapminder)
```

I have a very bad headache now.

