---
title: "HW05 Part 2: Importing"
author: "Molly Brakhane"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r}
library(tidyverse)
library(ggplot2)
```
# Limpets
```{r}
limpet <- read.csv("data/limpets.csv")
limpet 
```

```{r}
limpet$Site <- factor(limpet$Site, levels = c("PBL", "KNRM", "VBG", "WP", "PF", "DP", "SIO", "CTZ1", "CTZ2", "CNM"))
ggplot(data = limpet) +
  geom_boxplot(mapping = aes(x = Site, y = Length, color = Protected)) +
  labs(y = "Length (mm)")
```
Unprotected sites generally result in smaller limpet samples, with some outliers in WP and DP. Those that are protected are reliably larger.

# Roseate Terns
```{r}
roseate <- read.delim("data/roseate_terns.txt")
roseate_tidy <- roseate %>% 
  gather(`AR`, `AL`, `AC`, `BR`, `BL`, `BC`, `JR`, `JL`, `JC`, key = "Site", value = "Population")
roseate_tidy
filter(roseate_tidy, is.na(Population) == FALSE)
ggplot(data = roseate_tidy) +
  geom_line(mapping = aes(x = year, y = Population)) +
  facet_wrap(~Site, nrow = 3) +
  labs(x = "Year", y = "Population")
```
AL, AR, JL, and BL all obviously increased between 1992 and 2016, and JR and BR obviously decreased. Gaps in the population are results of a lack of populaation samples.

# Darters
```{r}
?read.fwf
start_pos <- c(1, 4, 15, 16, 17, 18, 21, 25, 29, 33, 36, 39, 40)
stop_pos <- c(3, 14, 15, 16, 17, 20, 24, 28, 32, 35, 38, 40, 41)
names <- c("Specimen ID", "Species", "Riffle", "Age", "Sex", "Length", "Depth", "Velocity", "Temperature", "MajSub", "MinSub", "MajType", "MinType")
darters <- read.fwf("data/darter.txt", skip = 15, widths = c(3, 11, 1, 1, 1, 3, 4, 3, 3, 2, 2, 1, 1),, col.names = names)
darters
```

```{r}
darter_riffle <- darters$riffle <- factor(darters$Riffle, ordered = FALSE, levels = c("1", "2"))
darters_majtype <- darters$majtype <- factor(darters$MajType, ordered = TRUE, levels = c( "s", "fg", "sg", "lg", "c"))
```
## Pipe Function
```{r}
f_darters <- darters %>% 
  filter(!Species %in% c("zonale", "tetrazonum")) %>% 
  select(-c(MinType, MinSub)) %>% 
  rename(Major_Type = MajType) %>% 
  rename(Major_Substrate = MajSub) %>% 
  arrange(Specimen.ID)

f_darters
```

## Plot 1
```{r}
ggplot(data = f_darters) +
  geom_point(mapping = aes(x = Length, y = Depth, color = Species, shape = Species)) +
  facet_wrap(.~Riffle, nrow = 3)
```
Caeruleum have the largest variance in Riffle 1, but Blennioides seem more suited to Riffle 2. In fact Blennioides grow almost twice as long in the second Riffle.
## Plot 2
```{r}
ggplot(data = f_darters) +
  geom_boxplot(mapping = aes(y = Length, x = Species)) +
  facet_wrap(~Riffle, nrow = 2)
```
Spectabile grow much much smaller in Riffle 2, whereas the three Blenioides I noticed in the previous graph are large outliers, so the Blenioides only grow slightly larger in riffle 2. Caeruleum grow significantly smaller in riffle 2 as well.
## Plot 3
```{r}
ggplot(data = f_darters) +
  geom_boxplot(mapping = aes(y = Length, x = Major_Substrate)) +
  facet_grid(Species~Riffle, nrow(3))
```
```{r}
ggplot(data = f_darters) +
  geom_boxplot(mapping = aes(y = Length, x = Major_Substrate)) +
  facet_grid(Riffle~Species, nrow(3))
```
This is a much easier to read graph than the one above. Spectabile show no major substrate preference in Riffle 2, while Blennioides and caeruleum show opposing changes in preference: where Blennoides move shallower, the caeruleum move deeper. It would appear shallower substrate is more beneficial for growth, as the shallower riffle shows longer individuals.
# Blacklip Abalone
## Chunk 1
```{r}
abalone <- read.csv("data/abalone.csv")
abalone
abalone1 <- select(abalone, -c(X))
ggplot(data = abalone1) +
  geom_boxplot(mapping = aes(x = Type, y = Height))
```
Immature are slightly smaller than the other two, but there is a significant outlier in the females.
## Chunk 2
```{r}
ggplot(data = abalone1) +
  geom_point(mapping = aes(x = Rings, y = Height))
```
No, they are not old. they are in fact younger than a lot of samples.
## Chunk 3
```{r}
abalone2 <- filter(abalone1, Height < 0.3)
abalone2$Type <- factor(abalone2$Type, ordered = TRUE, levels = c("I", "F", "M"))
ggplot(data = abalone2) +
  geom_point(mapping = aes(x = Rings, y = Height, color = Type))
```
There are some immatures that measure at a 0 height. Males and females appear equally spread out, thought there are some very small females and one small male. Most seem to become matture around 8 rings.
## Chunk 4
```{r}
abalone_I <- filter(abalone2, Height == 0)
abalone_I
is.na(abalone2) <- !abalone2
ggplot(data = abalone2) +
  geom_point(mapping = aes(x = Rings, y = Height, color = Type))
```
Again, most samples are mature after 8 rings. There are some immatures that have many rings and a large size though.
## Chunk 5
```{r}
ggplot(data = abalone2) +
  geom_point(mapping = aes(x = ShellWeight, y = LongestShell, color = Type))
ggplot(data = abalone2) +
  geom_point(mapping = aes(x = Rings, y = ShellWeight, color = Type))
```
Females seem to have very heavy shells, and there is a definite relatioship between rings and shell weight, and an even more significant relationship between shell weight and the largest shells, to no one's surprise.
