---
title: "HW 08 Part A: Play Date"
author: "Molly Brakhane"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---
# Installation
```{r}
library(tidyverse)
library(lubridate)
library(ggplot2)

```


# Part 1: Play Date

```{r}
birth_date <- "July 8, 1997"
anderson <- "9 June 1836"
anning <- "1799/05/21"
ball <- "July 24, 1892"
earle <- "August, 30 (1935)"
franklin <- "25th of July, 1920"
lederberg <- "December 18, 1922"
mcclintock <- "16061902"
mlk_birth <- "1/15/29"
mlk_dream <- "1963 August 08"
```

### Convert each date to POSIXct format, using lubridate functions, saving each of them to a unique variable. Print each result.

```{r}
birth_date <- mdy("July 8, 1997")
anderson <- dmy("9 June 1836")
anning <- ymd("1799/05/21")
ball <- mdy("July 24, 1892")
earle <- mdy("August, 30 (1935)")
franklin <- dmy("25th of July, 1920")
lederberg <- mdy("December 18, 1922")
mcclintock <- dmy("16061902")
mlk_birth <- mdy("1/15/29")
mlk_dream <- ymd("1963 August 08")

anderson_dttm <- as_datetime(anderson)
anning_dttm <- as_datetime(anning)
ball_dttm <- as_datetime(ball)
brakhane_dttm <- as_datetime(birth_date)
earle_dttm <- as_datetime(earle)
franklin_dttm <- as_datetime(franklin)
lederberg_dttm <- as_datetime(lederberg)
mcclintock_dttm <- as_datetime(mcclintock)
mlk_birth_dttm <- as_datetime(mlk_birth)
mlk_dream_dttm <- as_datetime(mlk_dream)

anderson_dttm
anning_dttm
ball_dttm
brakhane_dttm
earle_dttm
franklin_dttm
lederberg_dttm
mcclintock_dttm
mlk_birth_dttm
mlk_dream_dttm

```

### Check your dates to be sure they are correct. If you spot any errors, correct them now using the proper lubridate functions.

```{r}
mlk_dream_dttm_1 <- update(mlk_dream_dttm, year = 1963, month = 8, day = 28)
mlk_birth_dttm_1 <- update(mlk_birth_dttm, year = 1929, month = 1, day = 15)
```

### What day of the week were you born on? Show the full day name, not the abbreviated name or the day number.

```{r}
wday(brakhane_dttm, label = TRUE, abbr = FALSE)
```

### what day of the week would your birthday be on *this* year? Use lubridate's **update** function. You can hard code this year by you must figure out how to do it entirely by code for full points. In other words, your code should work just as well next year or five years in the future without editing.

```{r}
now <- now()
year <- year(now)
year
new_brakhane <- update(brakhane_dttm, year = year)
new_brakhane
wday(new_brakhane, label = TRUE, abbr = FALSE)
```

### What day of the week would your birthday be on in the Year 2525, if Man is still alive? How old would you be? You may hard code 2525. Show the abbreviated day of the week.
 
```{r}
future_brakhane <- update(brakhane_dttm, year = 2525, month = 7, day = 8)
wday(future_brakhane, label = TRUE, abbr = TRUE)
future_brakhane
brakhane_dttm
future_age <- interval(ymd(brakhane_dttm), ymd(future_brakhane))
time_length(future_age, "year")
```

### How many days was it between the birth of Dr. Elizabeth Garett Anderson and Alice Augusta Ball?

```{r}
days_between <- interval(ymd(anderson_dttm), ymd(ball_dttm))
days_between
time_length(days_between, unit = "day")
```

### What is the duration between MLK's birth date and his *I Had a Dream* speech?

```{r}
mlk <- interval(ymd(mlk_birth_dttm_1), ymd(mlk_dream_dttm_1))
mlk
time_length(mlk, "year")
```

## Enter the eight dttm dates into a vector, including your own, in the order they are given above. Name each element of the vector with the last name of each person.

```{r}
birth_vector <- c(brakhane_dttm, anderson_dttm, anning_dttm, ball_dttm, earle_dttm, franklin_dttm, lederberg_dttm, mcclintock_dttm, mlk_birth_dttm_1)
names(birth_vector) <- c("Brakhane", "Anderson", "Anning", "Ball", "Earle", "Franklin", "Lederberg", "McClintock", "King")
```

### 1. Of the birthdates, which of them are leap years? Start with POSIXct dates, display only the final years of the leap years, not the date. This should be three=four years.

```{r}
leaps <- leap_year(birth_vector)
leaps
year(birth_vector[leaps])

```



### 2. Subtract each birthdate in the vector from your own to calculate the number of days that have passed between their birth dates and yours. Sort from the greatest number to fewest. You should listed last.

```{r}
days_passed_a <- time_length(birth_vector - brakhane_dttm, "day")
sort(days_passed_a, decreasing = FALSE)
#JUST TO CHECK:
days_passed_b <- interval(ymd(mcclintock_dttm), ymd(brakhane_dttm))
time_length(days_passed_b, "day")
```


# Part 2

## Lake Ice
### Import lake_ice.csv
```{r}
library(readr)
library(dplyr)
library(tidyverse)
library(readr)
lake_ice <- read_csv("Data/lake_ice.csv")
lake_ice
```

### Filter the data to save only Lake Mendota.
```{r}
lake_mendota <- filter(lake_ice, lakename == "LAKE MENDOTA")
lake_mendota
```

### Select the columns (iceon_year, iceon_month, iceon_day, iceoff_year, iceoff_month, iceoff_day)

```{r}
select(lake_mendota, iceon_year, iceon_month, iceon_day, iceoff_year, iceoff_month, iceoff_day)
```



### Use lubridate to assemble the ice on year month and date into an ice_on date column. Then do it for an ice_off column.

```{r}
lake_mendota1 <- lake_mendota %>% 
  select(iceon_year, iceon_month, iceon_day, iceoff_year, iceoff_month, iceoff_day) %>% 
  mutate(ice_on = make_date(iceon_year, iceon_month, iceon_day),
         ice_off = make_date(iceoff_year, iceoff_month, iceoff_day))
```


### Create a new *duration* column by subtracting ice_on from ice_off. 

```{r}
lake_mendota2 <- lake_mendota1 %>% 
  mutate(duration = as.numeric(ice_off - ice_on))
lake_mendota2
```


### Make a line graph to show duration changed over iceoff_year. Use na.rm=TRUE. 

```{r}
ggplot(data = lake_mendota2) +
  geom_line(mapping = aes(x = iceoff_year, y = duration), na.rm = TRUE)+
  coord_cartesian(xlim = c(1850, 2030))
```


### Make appropriate labels.

```{r}
ggplot(data = lake_mendota2) +
  geom_line(mapping = aes(x = iceoff_year, y = duration), na.rm = TRUE)+
  coord_cartesian(xlim = c(1850, 2015)) +
  labs(x = "Year", y = "Ice Duration (days)")
```

## Soil Carbon Dioxide
### Import soil_co2.txt
```{r}
library(tidyverse)
library(lubridate)
soil_co2 <- read.delim("~/Masters/GABD/brakhane_molly/Data/soil_co2.txt")
soil_co2
```

### Establish a POSIXct-compatible date variable for 01 January 2006.

```{r}
jan <- make_date(year = 2006, day = 1, month = 1)
```


### Rename the columns so that they are syntactic.

```{r}
soil <- soil_co2 %>% 
  rename(Day = DayN..) %>% 
  rename(Date_sampled = Datecode..) %>% 
  rename(Minimum = Min..) %>% 
  rename(Vegetation = Veg..) %>% 
  rename(Plot = Plot..) %>% 
  rename(Group = Group..) %>% 
  rename(CO2_concentration = CorrCO2..) %>% 
  rename(Soil_type = SoilT..)
```


### Make a **Date** column that adds the day of CO2 measurement to the 01 Jan 2006 reference date. Create a **Year** column that extracts just the year. Create a **Month** colummnn that extracts just the month from the new **Date** coluumn as a three letter abreviation. Group data by Year and Month.

```{r}
soil1 <- soil %>% 
  mutate(Date = jan + Day,
         Year = year(Date),
         Month = month(Date, label = TRUE, abbr = TRUE)) %>% 
  group_by(Year, Month)
soil1
```


### Summarize mean CO2 and standard error of the mean. You had to calculate standard error for a previous assignment so you can review your previous code.

```{r}
soil2 <- soil1 %>% 
  summarise(
    meanCO2 = mean(CO2_concentration),
    se_Co2 = sd(CO2_concentration)/sqrt(NROW(soil1))
  )
soil2
```


### Make a column chart of mean CO2 for each month. Use either geom_errorbar or geom_linerange to plot the standard error of the mean. Use facet_wrap on Year to make plots the results by year. If you use geom_errorbar, then use the width argument to make reasonable horizontal lines.

```{r}
ggplot(data = soil2) +
  geom_col(aes(x = Month, y = meanCO2)) +
  geom_linerange(aes(x = Month, ymin = meanCO2 - se_Co2, ymax = meanCO2 + se_Co2)) +
  facet_wrap(~Year, nrow = 2)
```


### Make the same plot but use points instead of columns, and use only geom_linerange.

```{r}
ggplot(data = soil2) +
  geom_point(aes(x = Month, y = meanCO2)) +
  geom_linerange(aes(x = Month, ymin = meanCO2 - se_Co2, ymax = meanCO2 + se_Co2)) +
  facet_wrap(~Year, nrow = 2)
```



## Mississippi River Discharge
### Import and Wrangle
```{r}
library(tidyverse)
library(dplyr)

thebes <- read_csv("~/Masters/GABD/brakhane_molly/Data/thebes_discharge.csv")
thebes
colnames(thebes)
thebes1 <- thebes %>%
  gather ("1988", "1989", "1990", "1991", "1992", "1993", "1994", "1995", "1996", "1997", "1998", "1999", "2000", "2001", "2002", "2003","2004", "2005", "2006", "2007", "2008", "2009", "2010", "2011", "2012", "2013", "2014", "2015", "2016", "2017", "2018", key = "Year", value = "discharge") %>% 
  group_by(Year, Month) %>% 
  mutate(discharge = discharge / 1000,
         Event = case_when(
           Year== "1993" ~ "Flood",
           Year == "2011" ~ "Flood",
           Year == "1988" ~ "Drought",
           Year == "2012" ~ "Drought",
           TRUE ~ "Normal"),
         average_discharge = mean(discharge), na.rm = FALSE)
thebes1
```

### Make a second data set

```{r}
events <- thebes1 %>% 
  filter(Event %in% c("Flood", "Drought"))
events
```


### Plot the data

```{r}
p1 <- thebes1 %>%
  filter(!Year %in% c("1988", "1993", "2011", "2012")) %>%
  ggplot() +
  geom_line(aes(x = Month, y = average_discharge, group = Year),
            color = "gray70",
            size = 0.3) +
  labs(x = NULL,
       y = "Discharge (cf/s, thousands)")
p1
```


### We will now add the second layer using the smaller data set., by using this code format. Mapping and grouping is the same as above, but add a color = Event argument to the aes() mapping function.

```{r}
thebes1 %>%
  filter(!Year %in% c("1988", "1993", "2011", "2012")) %>%
  ggplot() +
  geom_line(aes(x = Month, y = average_discharge, group = Year),
            color = "gray70",
            size = 0.3) +
  labs(x = NULL,
       y = "Discharge (cf/s, thousands)") +
  geom_line(data = events, aes(x = Month, y = average_discharge, group = Year, color = Event))
```

