---
title: "HW 06 Part 1: Complete the sections"
author: "Molly Brakhane"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

- Change "your name" in the YAML header above to your name.

- As usual, enter the examples in code chunks and run them, unless told otherwise.

## Chapter 5: Data transformation

Read [R4ds Chapter 5: Data transformation](https://r4ds.had.co.nz/transform.html), sections 5-7.

You do not answer any of the questions this week. Just read and run all of the examples.
```{r}
library(nycflights13)
library(tidyverse)
```
```{r}
flights
```

```{r}
filter(flights, month == 1, day == 1)
```
Selecting only New Years Day flights.
```{r}
jan1 <- filter(flights, month == 1, day ==1)
(dec25 <- filter(flights, month == 12, day == 25))
```
Selecting only Christmas and New Years Day flights.
```{r}
filter(flights, month = 1)
```
```{r}
sqrt(2) ^ 2 == 2
```
```{r}
near(sqrt(2) ^ 2, 2)
near(1 / 49 * 49, 1)
```
```{r}
filter(flights, month == 11 | month == 12)
nov_dec <- filter(flights, month %in% c(11,12))
```
Filter for flights in November and December only
```{r}
filter(flights, !(arr_delay > 120 | dep_delay > 120))
filter(flights, arr_delay <= 120, dep_delay <= 120)
```
Flights that were neither 2 hours early or late.
```{r}
NA > 5
10 == NA
NA + 10
NA / 2
NA == NA
```
```{r}
x <- NA
y <- NA
x == y
```
```{r}
is.na(x)
```
```{r}
df <- tibble(x = c(1, NA, 3))
filter(df, x > 1)
filter(df, is.na(x) | x > 1)
```
```{r}
arrange(flights, year, month, day)
```
```{r}
arrange(flights, desc(dep_delay))
```
```{r}
df <- tibble(x = c(5, 2, NA))
arrange(df, x)
arrange(df, desc(x))
```
Shows how tibbles change their order based on your command.
```{r}
select(flights, year, month, day)
select(flights, year:day)
select(flights, -(year:day))
```
Shows how they change their display based on how you phrase the same thing.
```{r}
rename(flights, tail_num = tailnum)
```
Renaming columns
```{r}
select(flights, time_hour, air_time, everything())
```
rearranging the order of the table

### Section 5.5: Add new variables with `mutate()`

Type and run all of the examples.

Pay careful attention to the useful creation functions in section 5.5.1. 
```{r}
flights_sml <- select(flights,
                      year:day,
                      ends_with("delay"),
                      distance, 
                      air_time
                      )
mutate(flights_sml,
       gain = dep_delay - arr_delay,
       speed = distance / air_time * 60)
```
using mutate to add columns that are results of calculations on already-existing columns
```{r}
mutate(flights_sml,
       gain = dep_delay - arr_delay,
       hours = air_time / 60,
       gain_per_hour = gain / hours
       )
```
More of using mutate to add columns
```{r}
transmute(flights,
          gain = dep_delay - arr_delay,
          hours = air_time / 60,
          gain_per_hour = gain/hours
          )
```
transmute takes the columns you make to make a new df of just those
```{r}
transmute(flights,
  dep_time,
  hour = dep_time %/% 100,
  minute = dep_time %% 100
)
```
Using these new colums to find more specific data.

```{r}
(x <- 1:10)
lag(x)
lead(x)
```
Orders them in opposing orders based on lagging or leading

```{r}
x
cumsum(x)
cummean(x)
```
Totals the sum of the data or the mean of the data

```{r}
y <- c(1, 2, 2, NA, 3, 4)
min_rank(y)
min_rank(desc(y))
```
This is another way to order things based on values, like lag and lead, but pointing out that 'desc' reverses the order.

```{r}
row_number(y)
dense_rank(y)
percent_rank(y)
cume_dist(y)
```
Varying ways of ordering data.
### Section 5.6: Grouped summaries with `summarise()`

Read and run all of the examples. Pay careful attention to the useful summary functions. Unless you really like baseball, skip the baseball part of 5.6.3.

```{r}
summarise(flights, delay = mean(dep_delay, na.rm = TRUE))
```
summarizing the average delay of all flights

```{r}
by_day <- group_by(flights, year, month, day)
summarise(by_day, delay = mean(dep_delay, na.rm = TRUE))
```
Group by specifies the relevant info that summarise can then use to make lots of average delays
```{r}
by_dest <- group_by(flights, dest)
delay <- summarise(by_dest,
  count = n(),
  dist = mean(distance, na.rm = TRUE),
  delay = mean(arr_delay, na.rm = TRUE)
)
delay <- filter(delay, count > 20, dest != "HNL")
ggplot(data = delay, mapping = aes(x = dist, y = delay)) +
  geom_point(aes(size = count), alpha = 1/3) +
  geom_smooth(se = FALSE)
```
Looking how the distance affects the delay times with destination
```{r}
delays <- flights %>% 
  group_by(dest) %>% 
  summarise(
    count = n(),
    dist = mean(distance, na.rm = TRUE),
    delay = mean(arr_delay, na.rm = TRUE)
  ) %>% 
  filter(count > 20, dest != "HNL")
```
specifying data for graphing
```{r}
flights %>% 
  group_by(year, month, day) %>% 
  summarise(mean = mean(dep_delay))
```
If we don't put in the na.rm code, we won't get the accurate data. it'll give back NA
```{r}
flights %>% 
  group_by(year, month, day) %>% 
  summarise(mean = mean(dep_delay, na.rm = TRUE))
```
This gives only full data instead of a bunch of NA
```{r}
not_cancelled <- flights %>% 
  filter(!is.na(dep_delay), !is.na(arr_delay))
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(mean = mean(dep_delay))
```
Data ignoring missing flights
```{r}
delays <- not_cancelled %>% 
  group_by(tailnum) %>% 
  summarise(
    delay = mean(arr_delay)
  )
ggplot(data = delays, mapping = aes(x = delay)) + 
  geom_freqpoly(binwidth = 10)
```
how many flights are delayed by how much
```{r}
delays <- not_cancelled %>% 
  group_by(tailnum) %>% 
  summarise(
    delay = mean(arr_delay, na.rm = TRUE),
    n = n()
  )
ggplot(data = delays, mapping = aes(x = n, y = delay)) + 
  geom_point(alpha = 1/10)
```
Plotting the number of flights with certain delay times
```{r}
delays %>% 
  filter(n > 25) %>% 
  ggplot(mapping = aes(x = n, y = delay)) + 
    geom_point(alpha = 1/10)
```
Specifying data that is important and ignoring outliers
```{r}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(
    avg_delay1 = mean(arr_delay),
    avg_delay2 = mean(arr_delay[arr_delay > 0])
  )
```
calculating average arrival time, one ignoring those that left early
```{r}
not_cancelled %>% 
  group_by(dest) %>% 
  summarise(distance_sd = sd(distance)) %>% 
  arrange(desc(distance_sd))
```
finding the standard deviation using sd, and the ordering it in decreasing order
 
```{r}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(
    first = min(dep_time),
    last = max(dep_time)
  )
```
Finding the first plane that isn't cancelled and last plance that wasn't cancelled

```{r}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(
    first_dep = first(dep_time), 
    last_dep = last(dep_time)
  )
```
Finding the first and last departure of each day
```{r}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  mutate(r = min_rank(desc(dep_time))) %>% 
  filter(r %in% range(r))
```
Making the new variable r
```{r}
not_cancelled %>% 
  group_by(dest) %>% 
  summarise(carriers = n_distinct(carrier)) %>% 
  arrange(desc(carriers))
```
Listing destination in descending order by carrier
```{r}
not_cancelled %>% 
  count(dest)
```
How many planes are going to each destination
```{r}
not_cancelled %>% 
  count(tailnum, wt = distance)
```
showing miles flown by each plane
```{r}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(n_early = sum(dep_time < 500))
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(hour_perc = mean(arr_delay > 60))
```
Labeling the variables n early and hour perc
```{r}
daily <- group_by(flights, year, month, day)
(per_day   <- summarise(daily, flights = n()))
(per_month <- summarise(per_day, flights = sum(flights)))
(per_year  <- summarise(per_month, flights = sum(flights)))
```
narrowing the data to one category instead

```{r}
daily %>% 
  ungroup() %>%       
  summarise(flights = n())
```
ungrouping data
### Section 5.7: Grouped mutates (and filters)

Read and run all of the examples.
```{r}
flights_sml %>% 
  group_by(year, month, day) %>%
  filter(rank(desc(arr_delay)) < 10)
```
setting the new order
```{r}
popular_dests <- flights %>% 
  group_by(dest) %>% 
  filter(n() > 365)
popular_dests
```
finding popular destination
```{r}
popular_dests %>% 
  filter(arr_delay > 0) %>% 
  mutate(prop_delay = arr_delay / sum(arr_delay)) %>% 
  select(year:day, dest, arr_delay, prop_delay)
```
making prop delay